The Astropy Project: \\ Sustaining and Growing a Community-oriented Open-source Project and the Latest Major Release (v5.0) of the Core Package

### 2022-01-25 src/ms.tex
wrap at 80 chars

\begin{abstract}
The \astropy Project supports and fosters the development of open-source and openly-developed
\python packages that provide commonly-needed functionality to the astronomical
community.
A key element of the \astropy Project is the core package \astropypkg, which serves as the
foundation for more specialized projects and packages.
In this article, we summarize key features in the core package as of the recent major
release, version 5.0, and provide major updates for the project.
We then discuss supporting a broader ecosystem of inter-operable packages,
including connections with several astronomical observatories and missions.
We also revisit the future outlook of the \astropy Project and the current
status of Learn Astropy. 
We conclude by raising and discussing the current and future challenges facing the project.
\end{abstract}

@@

All affiliated packages are listed on the astropy 
website\footnote{\url{https://www.astropy.org/affiliated}}. Since 
\cite{astropy:2018}, there has been an explosion of gravitational-wave affiliated 
packages including: \texttt{PyCBC} for exploring GW sources, lenstronomy for 
modeling strong gravitational lenses, \texttt{ligo.skymap} for visualizing GW 
probability maps, and \texttt{EinsteinPy} for general relativity and gravitational astronomy. 
There have been several packages related to HEALPix: \texttt{astropy-healpix}
for a BSD licensed HEALPix implementation, and \texttt{mocpy} for Multi-Order 
Coverage maps. \texttt{astroalign} as been introduced for astrometric registration, 
and \texttt{python-cpl} has been added for ESO pipelines and VLT data products. 
For ground-based astronomy, \texttt{baseband} has added IO capabilities for 
VLBI, and \texttt{SpectraPy} brings slit spectroscopy to the astropy ecosystem. 
We have \texttt{agnpy} for AGN astrophysics, and \texttt{statmorph} for fitting galactic 
morphological diagnostics. \texttt{saba} gives an interface to sherpa's fitting 
routines. For modeling interstellar dust extinction we have 
\texttt{dust_extinction}, and for extracting features from time series we 
have \texttt{feets}. We also have \texttt{corral} for managing data intensive parallel 
pipelines. \texttt{BayesicFitting} provides an interface for generic Bayesian 
inference, and \texttt{sbpy} does calculations for asteroid and cometary 
astrophysics. Finally, \texttt{synphot} provides an interface for synthetic photometry.

Several of the affiliated packages described in [reference to paper 2] have
had substantial improvements. \texttt{astroquery} \cite{astroquery} has added
access to roughly a dozen new missions and data services, including JWST, and
the project has switched to a continuous release model. Every time a change
is committed to the main development branch it is published on the Python
Package Index (PyPI, ref) and available for installation. Formal releases are
still done a couple of times per year. \texttt{photutils} \cite
{photutils} released its first stable version, indicating that the API will
change less frequently, and there have been several significant performance
improvements.  \texttt{ccdproc} \cite{ccdproc} also released a new major
version, bringing better performance to some image combination operations.
The \texttt{regions} package \cite{pyregions}, for manipulating ds9-style
region definitions \cite{ds9}, added new ways to manipulate regions and
introduced new region types. \texttt{reproject}’s \cite{reproject} major new
feature is a function to align and co-add images to create a mosaic; better
support for parallelization was also added. \texttt{specutils}, the package
that defines containers for 1D and 2D spectra \cite{specutils}, also had its
first stable release and the addition of classes to read JWST data.

>>>>

\begin{abstract}

    To be written...

\end{abstract}

@@
